Looking for CDI Bias

### 2021-07-07 01-preliminary.Rmd
fitted more sex DIF models

---
title: "Looking for CDI Bias"
author: "Nathan, George, and Mike"
date: "`Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

@@

library(wordbankr)
library(tidyverse)
library(ggrepel)
library(mirt) 
library(here)
library(kableExtra)
library(gridExtra)
library(printr)
#library(parallel)
#library(doParallel)
#library(mclust)
#library(Rtsne)
#library(factoextra)
```

@@

```{r}
m1 <- mirt(d_mat, 1, "2PL", 
           technical=list(NCYCLES=500))

@@

```{r}
# https://rdrr.io/cran/mirt/man/DIF.html
# https://rdrr.io/cran/mirt/man/multipleGroup.html

@@

save(file = here("data/eng_wg_sex_slopes_invariant.Rds"), mg2)

Mit = as_tibble(coef(mg2, simplify=T)$Male$items) %>%
  mutate(definition = rownames(coef(mg2, simplify=T)$Male$items)) %>%
  select(-g, -u) %>% 
  rename(d_m = d)
Fit = as_tibble(coef(mg2, simplify=T)$Female$items) %>%
  mutate(definition = rownames(coef(mg2, simplify=T)$Female$items)) %>%
  select(-g, -u) %>%
  rename(d_f = d)

combo <- Mit %>% left_join(Fit) %>%
  mutate(d_diff = d_f - d_m,
         d_diff_abs = abs(d_diff))

thresh = median(combo$d_diff_abs) + sd(combo$d_diff_abs)

anova(m1, mg2) # AIC prefers group model, BIC prefers 2PL
mg2s <- summary(mg2)
fem = mg2s[[1]]$rotF
mal = mg2s[[2]]$rotF

plot(x=fem, y=mal) # slopes are constrained to be the same...
# where are the difficulties? 
```

>>>>

---
title: "Looking for CDI Bias"
author: "Nathan, George, and Mike"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

@@

library(wordbankr)
library(tidyverse)
library(ggrepel)
library(mirt) 
library(here)
#library(kableExtra)
library(gridExtra)
library(printr)
#library(parallel)
#library(doParallel)
#library(mclust)
#library(Rtsne)
#library(factoextra)
```

@@

```{r, eval=F}
m1 <- mirt(d_mat, 1, "2PL", 
           technical=list(NCYCLES=500))

@@

```{r, eval=F}
# https://rdrr.io/cran/mirt/man/DIF.html
# https://rdrr.io/cran/mirt/man/multipleGroup.html

@@

save(file = here("data/eng_wg_sex_slopes_invariant.Rds"), 
     mg2, m_sc2, mod_sc1, mod_fullconstrain)
```

```{r}
load(here("data/eng_wg_sex_slopes_invariant.Rds"))
load(here("data/eng_wg_mod_2pl.Rds"))

extract_group_df <- function(group_model) {
  Mit = as_tibble(coef(group_model, simplify=T)$Male$items) %>%
    mutate(definition = rownames(coef(group_model, simplify=T)$Male$items)) %>%
    select(-g, -u) %>% 
    rename(d_m = d)
  Fit = as_tibble(coef(group_model, simplify=T)$Female$items) %>%
    mutate(definition = rownames(coef(group_model, simplify=T)$Female$items)) %>%
    select(-g, -u) %>%
    rename(d_f = d)

  combo <- Mit %>% left_join(Fit) %>%
    mutate(d_diff = d_f - d_m,
           d_diff_abs = abs(d_diff)) 
  return(combo)
}

mg2_df = extract_group_df(mg2)

# same difficulties:
mod_sc1_df = extract_group_df(mod_sc1)

#anova(m1, mg2) # AIC prefers group model, BIC prefers 2PL
```

### Items with a large difference in difficulty across sex

```{r}
# items showing > median + 1SD of absolute difference in difficulty have a lot of DIF
thresh = median(mg2_df$d_diff_abs) + sd(mg2_df$d_diff_abs)
big_dif = mg2_df[which(mg2_df$d_diff_abs > thresh),]

#mg2s <- summary(mg2)
#fem = mg2s[[1]]$rotF
#mal = mg2s[[2]]$rotF
big_dif %>% arrange(desc(d_diff_abs)) #%>% kable()
```

### Plot item ease for males vs. females

We label the `r nrow(big_dif)` items with absolute ease difference of more than median + 1SD = `r round(thresh,2)`.

```{r, fig.width=8, fig.height=6}
ggplot(mg2_df, 
       aes(x = d_m, y = d_f)) + 
  geom_point(alpha=.7) + 
  ggrepel::geom_label_repel(data = filter(mg2_df, d_diff_abs > thresh), 
                            aes(label = definition), max.overlaps = 20) + theme_bw() +
  xlab("Item Ease for Males") + 
  ylab("Item Ease for Females") + 
  geom_abline(intercept=0, slope=1, linetype='dashed')
```
